---
title: "Supporting information 2: Integrated single-visit occupancy models"
subtitle: "A simulation study"
author: "Valentin Lauret, Hélène Labach, Matthieu Authier, Olivier Gimenez"
csl: apa.csl
output:
  html_document:
    code_folding: hide
    highlight: textmate
    theme: readable
    toc: yes
link-citations: yes
bibliography: SVbib.bib
---

<style>
p.caption {
  font-style: italic;
}
</style>

Supporting information 2 of article _Using single visits into integrated occupancy models to make the most of existing monitoring programs_.  

Comparing single-visit and repeated-visits integrated occupancy models.

# Methods

Required packages 

```{r message=FALSE, warning=FALSE, cache=TRUE}
library(R2jags)
library(tidyverse)
library(cowplot)
```

## Data simulation 

### State process

The occupancy state $z$ was drawn from a Bernoulli distribution with parameter $\psi$, $z \sim dbern(psi)$. We wrote $\psi$ as a logistic function of an environmental covariate `cov`: 

$$\mbox{logit}(\psi) = \alpha_{0} + \alpha_{1} \; \mbox{cov}$$

where $\alpha_{0}$ and $\alpha_{1}$ are unknown parameters that need to be estimated.  

We considered two sets of values for the alpha’s:  
	
  - $\alpha_0 = -1$ and $\alpha_1 = 0.2$ that led to $\psi$ approx. equal to 0.1
  - $\alpha_0 = -0.5$ and $\alpha_1 = 0.2$ that led to $\psi$ approx. equal to 0.3

  
We used the `sim_state_cov()` function to simulate the $z$ vector that contains the latent ecological state.
 
 <center>
```{r echo=FALSE, message=FALSE, warning=FALSE, cache = TRUE}
nsites <- 400
covenv <- runif(nsites,0,1)
covparam.3 <- c(-0.5,0.2,0) # psi = 0.3
cov_rel.3 <- as.numeric(covparam.3[1]+covparam.3[2]*scale(covenv)+covparam.3[3]*scale(covenv)*scale(covenv))

covparam.1 <- c(-1.9,0.2,0) # psi = 0.1
cov_rel.1 <- as.numeric(covparam.1[1]+covparam.1[2]*scale(covenv)+covparam.1[3]*scale(covenv)*scale(covenv))

 ggplot() + geom_line(aes(x=covenv,y=1/(1+exp(-cov_rel.3)), color="High occupancy"), lwd= 2) +
   geom_line(aes(x=covenv,y=1/(1+exp(-cov_rel.1)),color="Low occupancy"), lwd= 2) +
   ylim(0,1) + ylab("Occupancy probability : Ψ") +xlab("Fictive covariate") + theme_bw(base_size = 14) + scale_color_viridis_d()
```
 </center>


### Observation process

The observations are drawn either from a Bernoulli distribution with parameter $p$ or a multinomial distribution depending on two detection probability $p_1$ and $p_2$ for the integrated occupancy model. We wrote $p$ as a logistic function of a sampling effort covariate $\mbox{seff}$:  

$$\mbox{logit}(p) = \beta_{0} + \beta_{1} \; \mbox{seff}$$

where $\beta_{0}$ and $\beta_{1}$ are unknown parameters that need to be estimated.   

We considered two sets of values for the beta’s:

  - $\beta_0 = -0.1$ and $\beta_1 = 0.16$ that led to $p$ approx. equal to 0.5
  - $\beta_0 = 1.8$ and $\beta_1 = 0.6$ that led to $p$ approx. equal to 0.8

We used the `sim_obs_cov()` function to simulate the $y$ vector that contains the detections/non-detections.


 <center>
```{r echo=FALSE, cache=TRUE}
 nsites = 400
Seff.5 <- data.frame(cbind(rnorm(nsites),rnorm(nsites),rnorm(nsites),rnorm(nsites),rnorm(nsites)))
obsparam.5 <- c(-0.1,0.16,0)

Seff.8 <- data.frame(cbind(rnorm(nsites),rnorm(nsites),rnorm(nsites),rnorm(nsites),rnorm(nsites),rnorm(nsites),rnorm(nsites),rnorm(nsites)))

obsparam.8 <- c(1.8,0.6,0)

obs_rel.5 <- as.numeric(obsparam.5[1]+obsparam.5[2]*Seff.5[,1]+obsparam.5[3]*Seff.5[,1]*Seff.5[,1])

obs_rel.8 <- as.numeric(obsparam.8[1]+obsparam.8[2]*Seff.8[,1]+obsparam.8[3]*Seff.8[,1]*Seff.8[,1])

 ggplot() + geom_line(aes(x=Seff.8[,1],y=1/(1+exp(-obs_rel.8)), color="High detection"), lwd= 2) +
   geom_line(aes(x=Seff.5[,1],y=1/(1+exp(-obs_rel.5)),color="Low detection"), lwd= 2) +
   ylim(0,1) + ylab("Detection probability : p") +xlab("Sampling effort covariate") + theme_minimal() + scale_color_viridis_d(begin = 0.2, end = 0.9)
```
 </center>

We simulated 4 different sampling occasions for repeated-visit occupancy. For single-visit occupancy, we formatted the data as described in the Methods section. For the sampling effort at site $i$, we summed the sampling effort over the 4 sampling occasions at site $i$, with `seff_sv <- apply(seff_rv,1,sum)`. For the detection/non-detection matrix, we considered the site as occupied if the species was detected at least once over the 4 sampling occasions, with `y_sv <- apply(y_rv,1,max)`.

### Study area 

We simulated $z$ and $y$ for study areas of either 100, or 400 sites. 

## Models

We fitted 4 different occupancy models:  

  - Repeated-visit (RV) single-method occupancy model (SOM) with low detection probability
  - Single-visit (SV) single-method occupancy model (SOM) with low detection probability
  - Repeated-visits integrated occupancy model (IOM RV), in which we combined two datasets into an integrated occupancy framework
  - Single-visit integrated occupancy model (IOM SV)

**Note that for the integrated occupancy, we combined datasets generated from the same $z$ ecological states with two different detection probabilities, i.e. we combined one monitoring with 0.5 detection probability (i.e. 'low $p$'), and one monitoring with 0.8 detection probability (i.e. 'high $p$').**

## Scenarios 

We considered 2 ecological scenarios depending on either the size of the study area or the value of occupancy probability:  

  - 100 sites with low occupancy, $\psi = 0.1$
  - 100 sites with high occupancy, $\psi = 0.3$
  
For each scenario, we ran 100 simulations and fitted the 4 occupancy models.


## Performances

To compare the performances of the different occupancy models we calculated the relative bias (RB), and the root mean square error (RMSE) over the `S = 100` simulations :  

  - Relative bias: $RB = \frac{1}{S} \sum_{1}^{S} \frac{(\widehat{\theta}_{s}-\theta)}{\theta}$  
  
  - Root Mean Square Error: $RMSE = \sqrt{ \frac{1}{S} \sum_{1}^{S} (\widehat{\theta}_{s}-\theta)^2}$  
  
## R codes

R functions to simulate the data and run the simulations  

```{r eval=FALSE}
# load functionS
sim_state_cov <- function(nb_sites=nsites,init_occ = covparam, covenv = covenv){
  
  # define various quantities
  R = nb_sites # number of sites
  
  l.psi <- init_occ[1] + init_occ[2] * covenv + init_occ[3] * covenv * covenv
  psi1 <-  1 / ( 1 + exp( -l.psi ))# quadratic relation of the covariate with initial occupancy
  
  # pre-allocate memory
  site <- 1:R # Sites
  
  z <- array(dim = c(R)) # Expected and realized occurrence
  
  # define state process
  # first year/season
  z <- rbinom(R, 1, psi1) # Initial occupancy state
  
  
  return(z)
}
sim_obs_cov <- function(z=z, nb_occ = nocc,Seff, obsparam = obsparam){
  
  # define various quantities
  R = length(z) # number of sites
  K = nb_occ # number of occasions
  
  y <- array(NA, dim = c(R,K))
  p <- array(NA, dim = c(R,K))
  
  # define observation process
  
  for(i in 1:R){
    for(k in 1:K){
      
      l.p <- obsparam[1] + obsparam[2] * Seff[i,k] + obsparam[3] * Seff[i,k] * Seff[i,k]# quadratic relation of the covariate with initial occupancy
      
      p[i,k] <- ( 1 / ( 1 + exp(- l.p )))
      
      prob <- z[i] * p[i,k]
      y[i,k] <- rbinom(1,1, prob= prob)
    } #k
  } #i
  
  
  # format data
  return(y)
}
loop <- function(nsites, nocc, nsim, covparam, obsparam1,obsparam2, Seff, Seff2, Covenv){
  
  
  #### allocate memories for posterior estimates ####
  
  p_i <- rep(NA, nsim)
  psi_i <- rep(NA, nsim)

  bcov_i <- rep(NA, nsim)
  bseff_i <- rep(NA, nsim)

  p_ii <- rep(NA, nsim)
  psi_ii <- rep(NA, nsim)

  bcov_ii <- rep(NA, nsim)
  bseff_ii<- rep(NA, nsim)

  p1_iii <- rep(NA, nsim)
  p2_iii <- rep(NA, nsim)
  psi_iii <-rep(NA, nsim)

  bcov_iii <- rep(NA, nsim)
  bseff1_iii<- rep(NA, nsim)
  bseff2_iii<- rep(NA, nsim)

  p1_iv <- rep(NA, nsim)
  p2_iv <- rep(NA, nsim)
  psi_iv <-rep(NA, nsim)

  bcov_iv <- rep(NA, nsim)
  bseff1_iv<- rep(NA, nsim)
  bseff2_iv<- rep(NA, nsim)

  # bis high p 
  p_ibis <- rep(NA, nsim)
  psi_ibis <- rep(NA, nsim)

  bcov_ibis <- rep(NA, nsim)
  bseff_ibis <- rep(NA, nsim)

  p_iibis <- rep(NA, nsim)
  psi_iibis <- rep(NA, nsim)

  bcov_iibis <- rep(NA, nsim)
  bseff_iibis<- rep(NA, nsim)

  # allocate memory to saved parameters
  
  rbpsi_i    <- rep(NA, nsim)
  rmsepsi_i   <- rep(NA, nsim)
  rbcov_i     <- rep(NA, nsim)
  rmsecov_i <- rep(NA, nsim)
  rbp_i     <- rep(NA, nsim)
  rmsep_i    <- rep(NA, nsim)

  rbseff_i    <- rep(NA, nsim)
  rmseseff_i  <- rep(NA, nsim)
  # ii<- rep(NA, nsim)
  rbpsi_ii    <- rep(NA, nsim)
  rmsepsi_ii  <- rep(NA, nsim)
  
  rbcov_ii    <- rep(NA, nsim)
  rmsecov_ii  <- rep(NA, nsim)
 
  rbp_ii      <- rep(NA, nsim)
  rmsep_ii    <- rep(NA, nsim)

  rbseff_ii   <- rep(NA, nsim)
  rmseseff_ii <- rep(NA, nsim)

  # iii   <- rep(NA, nsim)
  rbpsi_iii   <- rep(NA, nsim)
  rmsepsi_iii <- rep(NA, nsim)
  
  rbcov_iii   <- rep(NA, nsim)
  rmsecov_iii <- rep(NA, nsim)
  
  rbp1_iii    <- rep(NA, nsim)
  rmsep1_iii  <- rep(NA, nsim)
  rbseff1_iii <- rep(NA, nsim)
  rmseseff1_iii<- rep(NA, nsim)
  
  rbp2_iii    <- rep(NA, nsim)
  rmsep2_iii  <- rep(NA, nsim)
  rbseff2_iii <- rep(NA, nsim)
  rmseseff2_iii<- rep(NA, nsim)
                         # iv
  rbpsi_iv    <- rep(NA, nsim)
  rmsepsi_iv  <- rep(NA, nsim)
  rbcov_iv    <- rep(NA, nsim)
  rmsecov_iv  <- rep(NA, nsim)
  rbp1_iv     <- rep(NA, nsim)
  rmsep1_iv   <- rep(NA, nsim)
  rbseff1_iv  <- rep(NA, nsim)
  rmseseff1_iv<- rep(NA, nsim)
  rbp2_iv     <- rep(NA, nsim)
  rmsep2_iv   <- rep(NA, nsim)
  rbseff2_iv  <- rep(NA, nsim)
  rmseseff2_iv<- rep(NA, nsim)
  
  # bis highp    
  p_ibis <- rep(NA, nsim)
  psi_ibis <- rep(NA, nsim)

  bcov_ibis <- rep(NA, nsim)
  bseff_ibis <- rep(NA, nsim)

  p_iibis <- rep(NA, nsim)
  psi_iibis <- rep(NA, nsim)

  bcov_iibis <- rep(NA, nsim)
  bseff_iibis<- rep(NA, nsim)
  
  # parameters
  rbpsi_ibis    <- rep(NA, nsim)
  rmsepsi_ibis   <- rep(NA, nsim)
  rbcov_ibis     <- rep(NA, nsim)
  rmsecov_ibis <- rep(NA, nsim)
  rbp_ibis     <- rep(NA, nsim)
  rmsep_ibis    <- rep(NA, nsim)
  
  rbseff_ibis    <- rep(NA, nsim)
  rmseseff_ibis  <- rep(NA, nsim)
  # ii<- rep(NA, nsim)
  rbpsi_iibis    <- rep(NA, nsim)
  rmsepsi_iibis  <- rep(NA, nsim)
  
  rbcov_iibis    <- rep(NA, nsim)
  rmsecov_iibis  <- rep(NA, nsim)
  
  rbp_iibis      <- rep(NA, nsim)
  rmsep_iibis    <- rep(NA, nsim)
  
  rbseff_iibis   <- rep(NA, nsim)
  rmseseff_iibis <- rep(NA, nsim)

  # JAGS models 
  
  sink("Sim_occu_cov.jags")
  cat("
    model {
    
    # Specify priors
    
    alpha.psi ~ dnorm(0,0.444) # intercept occupancy
    alpha.p ~ dnorm(0,0.444) # intercept detection

    beta.cov ~ dnorm(0,0.444) # slope cov effect
    beta.qcov ~ dnorm(0,0.444)	# slope quadratic cov effect
    
    beta.seff ~ dnorm(0,0.444)	# slope Sampling effort effect
    beta.qseff ~ dnorm(0,0.444)	# slope quadratic sampling effort
    
    beta.occ2 ~ dnorm(0,0.444) # occasion effect
    beta.occ3 ~ dnorm(0,0.444) # occasion effect
    beta.occ4 ~ dnorm(0,0.444) # occasion effect
    
    
    ### Ecological states
    
    for (i in 1:nsite){
    
      z[i] ~ dbern(muZ[i])
      logit(muZ[i]) <- l.muZ[i]
      
      l.muZ[i] <- alpha.psi + beta.cov * covenv[i] + beta.qcov * covenv[i] * covenv[i]
    
    } # i
  
    # Observation model
   
    for (i in 1:nsite){
    
      for (k in 1:nocc){
    
          logit(p[i,k]) <- -l.p[i,k]
          l.p[i,k] <- alpha.p + beta.seff * Seff[i,k] + beta.qseff * Seff[i,k] * Seff[i,k] + beta.occ2 * equals(k,2) + beta.occ3 * equals(k,3) + beta.occ4 * equals(k,4)
          

    y[i,k] ~ dbern(z[i]*p[i,k])    
    
      } #k
    } #i
    
    
    }
    ",fill = TRUE)
  sink()
  
  sink("sim_IOM_cov.jags")
  cat("
    model {
    # priors

    alpha.psi ~ dnorm(0,0.444) # occupancy intercept
    alpha.p1 ~ dnorm(0,0.444) # detection samm intercept 
    alpha.p2 ~ dnorm(0,0.444) # detection gdegem intercept 
    
    # slope priors 
    
    beta.cov ~ dnorm(0,0.444) # slope cov effect
    beta.qcov ~ dnorm(0,0.444)	# slope quadratic cov effect
    
    beta.seff1 ~ dnorm(0,0.444)	# slope Sampling effort effect
    beta.qseff1 ~ dnorm(0,0.444)	# slope quadratic sampling effort
    
    beta.seff2 ~ dnorm(0,0.444)	# slope Sampling effort effect
    beta.qseff2 ~ dnorm(0,0.444)	# slope quadratic sampling effort
    
    beta.occ2 ~ dnorm(0,0.444) # occasion effect
    beta.occ3 ~ dnorm(0,0.444) # occasion effect
    beta.occ4 ~ dnorm(0,0.444) # occasion effect

    # NO COV
    
    # Ecological state 
    
    for (i in 1:nsite){
    
       z[i] ~ dbern(muZ[i])
      logit(muZ[i]) <- l.muZ[i]
      
        l.muZ[i] <- alpha.psi + beta.cov * covenv[i] + beta.qcov * covenv[i] * covenv[i]

      for (j in 1:nocc){

        # p1
        logit(p1[i,j]) <- lp1[i,j] 
        lp1[i,j] <- alpha.p1  + beta.seff1 * Seff1[i,j] + beta.seff1 * Seff1[i,j] * Seff1[i,j] + beta.occ2 * equals(j,2) + beta.occ3 * equals(j,3) + beta.occ4 * equals(j,4)
        
         # p2
        logit(p2[i,j]) <- lp2[i,j] 
        lp2[i,j] <- alpha.p2 + beta.seff2 * Seff2[i,j] + beta.qseff2 * Seff2[i,j] * Seff2[i,j] + beta.occ2 * equals(j,2) + beta.occ3 * equals(j,3) + beta.occ4 * equals(j,4)
          
        
        mu.p[i,j,1] <- 1 - z[i] + z[i] * (1 - p1[i,j]) * (1 - p2[i,j]) # P(O=0) : no detection
        mu.p[i,j,2] <- z[i] * p1[i,j] * (1 - p2[i,j])                  # P(O=1) : detection by samm only
        mu.p[i,j,3] <- z[i] * p2[i,j] * (1 - p1[i,j])                 # P(O=2) : detection by gdegem only
        mu.p[i,j,4] <- z[i] * p1[i,j] * p2[i,j]                        # P(O=3) : detection by both samm and gdegem

        y[i,j] ~ dcat(mu.p[i,j,])
    } #j
    } #i
    }
    ", fill = TRUE)
  sink()
  
  #### loop over nsim ####
  for( s in 1:nsim){
    
    # simulate datasets s
    
    z <- sim_state_cov(nb_sites = nsites, covenv= covenv, init_occ= covparam)
    y <- sim_obs_cov(z,nocc,Seff= Seff, obsparam= obsparam1)
    
    y1 <- y # the first detection dataset
    p1 <- obsparam1
    p2 <- obsparam2
    y2bis <- y2 <- sim_obs_cov(z,nocc,obsparam = obsparam2, Seff =Seff2) # another dataset with different p
    y2bis[y2>0] <- 2 
    y_int <- y1 + y2bis # 0 -> no detection, 1 -> detection by monitoring 1, 2-> detection by 2, 3 -> detection by both methods
    
    
    
    # perform analysis
    
    # Specify MCMC settings one for all
    ni <-  4000
    nb <- 1000
    nc <- 1
    
    #### i)  One monitoring RV low p ####
    
    
    # Bundle data
    
    win.data <- list(y = y1,nsite = dim(y)[1], nocc = dim(y)[2], Seff = Seff, covenv = covenv)
    
    # Initial values
    zst <- apply(y, 1, max)	# Observed occurrence as inits for z
    psi = covparam[1]    # sum(z)/nsites
    p = obsparam1[1]
    bcov <- covparam[2]
    bseff <- obsparam1[2]
    
    inits <- function(){ list(z = zst, alpha.psi = psi, alpha.p= p, beta.cov = bcov, beta.seff= bseff)}
    
    # Parameters monitored
    params <- c( "alpha.p","alpha.psi", "beta.cov", "beta.qcov","beta.seff" ,"beta.qseff") 
    
    
    # Call JAGS from R (BRT 3 min)
    out <- jags(win.data, inits, params, "Sim_occu_cov.jags", n.chains = nc, n.iter = ni, n.burnin = nb, working.directory = getwd())
    
    # denplot(out)
    # head(out$BUGSoutput$summary)
    
    # save results 
    
    # intercept
    p_i[s] <- out$BUGSoutput$mean$alpha.p
    psi_i[s] <- out$BUGSoutput$mean$alpha.psi

    # slope linear estimate (Q: do we have to save the quadratic estimates-)
    
    bcov_i[s] <-    out$BUGSoutput$mean$beta.cov
    bseff_i[s] <-   out$BUGSoutput$mean$beta.seff

    #### i bis )  One monitoring RV high p ####
    
    
    # Bundle data
    
    win.data <- list(y = y2,nsite = dim(y2)[1], nocc = dim(y2)[2], Seff = Seff2, covenv = covenv)
    
   
    # Initial values
    zst <- apply(y2, 1, max)	# Observed occurrence as inits for z
    psi = covparam[1]   # sum(z)/nsites
    p = obsparam2[1]
    bcov <- covparam[2]
    bseff <- obsparam2[2]
    
    inits <- function(){ list(z = zst, alpha.psi = psi, alpha.p= p, beta.cov = bcov, beta.seff= bseff)}
    
    # Parameters monitored
    params <- c( "alpha.p","alpha.psi", "beta.cov", "beta.qcov","beta.seff" ,"beta.qseff") 
    
    
    # Call JAGS from R (BRT 3 min)
    out <- jags(win.data, inits, params, "Sim_occu_cov.jags", n.chains = nc, n.iter = ni, n.burnin = nb, working.directory = getwd())
    
   # head(out$BUGSoutput$summary)
    
    # save results 
    
    # intercept
    p_ibis[s] <- out$BUGSoutput$mean$alpha.p
    psi_ibis[s] <- out$BUGSoutput$mean$alpha.psi

    # slope linear estimate (Q: do we have to save the quadratic estimates-)
    
    bcov_ibis[s] <- out$BUGSoutput$mean$beta.cov
    bseff_ibis[s] <- out$BUGSoutput$mean$beta.seff

    
     #### ii)  One monitoring SV ####
    
    # same model jags than i)
    
    # Bundle data
    
    ysv <-  matrix(apply(y,1,max))
    Seffsv <- matrix(apply(Seff,1,sum))
    
    win.data <- list(y = ysv,nsite = dim(ysv)[1], nocc = dim(ysv)[2], Seff = Seffsv, covenv = covenv)
    
    
    # Initial values
    zst <- apply(ysv, 1, max)	# Observed occurrence as inits for z
    psi = covparam[1]   # sum(z)/nsites
    p = obsparam1[1]
    bcov <- covparam[2]
    bseff <- obsparam1[2]
    
    inits <- function(){ list(z = zst, alpha.psi = psi, alpha.p= p, beta.cov = bcov, beta.seff= bseff)}
    
    
    # Parameters monitored
    params <- c("z", "alpha.p","alpha.psi", "beta.cov", "beta.seff","beta.qcov", "beta.qseff") 
    
    
    # Call JAGS from R (BRT 3 min)
   
    out <- jags(win.data, inits, params, "Sim_occu_cov.jags", n.chains = nc,  n.iter = ni,  n.burnin = nb, working.directory = getwd())
    
    # head(out$BUGSoutput$summary)
    
    p_ii[s] <- out$BUGSoutput$mean$alpha.p
    psi_ii[s] <- out$BUGSoutput$mean$alpha.psi

    # slope linear estimate (Q: do we have to save the quadratic estimates-)
    
    bcov_ii[s] <- out$BUGSoutput$mean$beta.cov
    bseff_ii[s] <- out$BUGSoutput$mean$beta.seff

    #### ii bis )  One monitoring SV high p ####
    
    # same model jags than i)
    
    # Bundle data
    
    y2sv <-  matrix(apply(y2,1,max))
    
    Seff2sv <- matrix(apply(Seff2,1,max))
    win.data <- list(y = y2sv,nsite = dim(y2sv)[1], nocc = dim(y2sv)[2], Seff = Seff2sv, covenv = covenv)
    
    
    # Initial values
    zst <- apply(y2sv, 1, max)	# Observed occurrence as inits for z
    psi = covparam[1]   # sum(z)/nsites
    p = obsparam2[1]
    bcov <- covparam[2]
    bseff <- obsparam2[2]
    
    inits <- function(){ list(z = zst, alpha.psi = psi, alpha.p= p, beta.cov = bcov, beta.seff= bseff)}
    
    
    
    # Parameters monitored
    params <- c("z", "alpha.p","alpha.psi", "beta.cov", "beta.seff") 
    
    
    # Call JAGS from R (BRT 3 min)
    out <- jags(win.data, inits, params, "Sim_occu_cov.jags", n.chains = nc,  n.iter = ni,  n.burnin = nb, working.directory = getwd())
    

    p_iibis[s] <- out$BUGSoutput$mean$alpha.p
    psi_iibis[s] <- out$BUGSoutput$mean$alpha.psi

    # slope linear estimate (Q: do we have to save the quadratic estimates-)
    
    bcov_iibis[s] <- out$BUGSoutput$mean$beta.cov
    bseff_iibis[s] <- out$BUGSoutput$mean$beta.seff

    
    #### iii)  Two monitoring RV ####
    
    # we work with y_int
    win.data <- list(y = y_int+1,nsite = dim(y_int)[1], nocc = dim(y_int)[2], covenv = covenv, Seff1 = Seff1, Seff2 = Seff2)
    

    # Initial values
    zst <- apply(y_int, 1, max)
    zst[zst>0] <- 1 # Observed occurrence as inits for z
    psi = covparam[1]   # sum(z)/nsites
    p1 = obsparam1[1]
    p2 = obsparam2[1]
    bcov <- covparam[2]
    bseff1 <- obsparam1[2]
    bseff2 <- obsparam2[2]
    
    inits <- function(){ list(z = zst, alpha.psi = psi, alpha.p1=p1, alpha.p2=p2,beta.cov = bcov, beta.seff1= bseff1, beta.seff2= bseff2)}
    
    # Parameters monitored
    params <- c("z", "alpha.p1","alpha.psi","alpha.p2","beta.cov","beta.qcov","beta.seff1","beta.seff2","beta.qseff1","beta.qseff2") 
    
    
    # Call JAGS from R (BRT 3 min)
    out <- jags(win.data, inits, params, "sim_IOM_cov.jags", n.chains = nc,  n.iter = ni,n.burnin = nb, working.directory = getwd())
    
    p1_iii[s] <-out$BUGSoutput$mean$alpha.p1
    p2_iii[s] <- out$BUGSoutput$mean$alpha.p2

    psi_iii[s] <- out$BUGSoutput$mean$alpha.psi

    # save results slope 
    
    bseff1_iii[s] <-out$BUGSoutput$mean$beta.seff1

    bseff2_iii[s] <- out$BUGSoutput$mean$beta.seff2

    bcov_iii[s] <- out$BUGSoutput$mean$beta.cov

    
    #### iv)  Two monitoring SV ####
    
    # same model jags than iii)
    
    # Bundle data
    
    y1sv <-  matrix(apply(y1,1,max))
    y2sv <-  matrix(apply(y2,1,max))
    y2sv[y2sv>0] <- 2
    
    ysv_int <- y1sv + y2sv
    
    Seff1sv <-  matrix(apply(Seff1,1,sum))
    Seff2sv <-  matrix(apply(Seff2,1,sum))
    
    win.data <- list(y = ysv_int+1,nsite = dim(ysv_int)[1], nocc = dim(ysv_int)[2], Seff1= Seff1sv, Seff2 = Seff2sv, covenv = covenv)
    
    # Initial values
    zst <- apply(ysv_int, 1, max)
    zst[zst>0] <- 1 # Observed occurrence as inits for z
    psi = covparam[1]   # sum(z)/nsites
    p1 = obsparam1[1]
    p2 = obsparam2[1]
    bcov <- covparam[2]
    bseff1 <- obsparam1[2]
    bseff2 <- obsparam2[2]
    
    inits <- function(){ list(z = zst, alpha.psi = psi, alpha.p1=p1, alpha.p2=p2,beta.cov = bcov, beta.seff1= bseff1, beta.seff2= bseff2)}
    
    # Parameters monitored
    params <- c("z", "alpha.p1","alpha.psi","alpha.p2","beta.cov","beta.qcov","beta.seff1","beta.seff2","beta.qseff1","beta.qseff2") 
    
    
    # Call JAGS from R (BRT 3 min)
    out <- jags(win.data, inits, params, "sim_IOM_cov.jags", n.chains = nc,  n.iter = ni, n.burnin = nb, working.directory = getwd())
    
    p1_iv[s] <-out$BUGSoutput$mean$alpha.p1
    p2_iv[s] <- out$BUGSoutput$mean$alpha.p2

    psi_iv[s] <-out$BUGSoutput$mean$alpha.psi

    # save results slope 
    
    bseff1_iv[s] <-out$BUGSoutput$mean$beta.seff1
    bseff2_iv[s] <- out$BUGSoutput$mean$beta.seff2
    bcov_iv[s] <- out$BUGSoutput$mean$beta.cov

    # calcul and save RMSE and RB
    
    # i 
    rbpsi_i[s]        <- (psi_i[s]-covparam[1])/covparam[1]   # RB alpha.psi
    rmsepsi_i[s]      <- (psi_i[s] - covparam[1])^2 # RMSE alpha.psi
    
    rbcov_i[s]        <- (bcov_i[s]-covparam[2])/covparam[2]   # RB cov
    rmsecov_i[s]      <- (bcov_i[s] - covparam[2])^2 # RMSE cov
    
    rbp_i[s]          <- (p_i[s]-obsparam1[1])/obsparam1[1]       # RB alpha p 
    rmsep_i[s]        <- (p_i[s] - obsparam1[1])^2 # RMSE alpha p
    
    rbseff_i[s]       <- (bseff_i[s]-obsparam1[2])/obsparam1[2]   # RB obs
    rmseseff_i[s]     <- (bseff_i[s] - obsparam1[2])^2 # RMSE obs
    
    
    # ii
    rbpsi_ii[s]       <- (psi_ii[s]-covparam[1])/covparam[1]   # RB alpha.psi
    rmsepsi_ii[s]     <- (psi_ii[s] - covparam[1])^2 # RMSE alpha.psi
    
    rbcov_ii[s]       <- (bcov_ii[s]-covparam[2])/covparam[2]  # RB cov
    rmsecov_ii[s]     <- (bcov_ii[s] - covparam[2])^2 # RMSE cov
    
    rbp_ii[s]         <- (p_ii[s]-obsparam1[1])/obsparam1[1]      # RB alpha p 
    rmsep_ii[s]       <- (p_ii[s] - obsparam1[1])^2 # RMSE alpha p
    
    rbseff_ii[s]      <- (bseff_ii[s]-obsparam1[2])/obsparam1[2]   # RB obs
    rmseseff_ii[s]    <- (bseff_ii[s] - obsparam1[2])^2 # RMSE obs
    
    
    # iii   
    rbpsi_iii[s]      <- (psi_iii[s]-covparam[1])/covparam[1]  # RB alpha.psi
    rmsepsi_iii[s]    <- (psi_iii[s] - covparam[1])^2 # RMSE alpha.psi
    
    rbcov_iii[s]      <- (bcov_iii[s]-covparam[2])/covparam[2]   # RB cov
    rmsecov_iii[s]    <- (bcov_iii[s] - covparam[2])^2 # RMSE cov
    
    rbp1_iii[s]       <- (p1_iii[s]-obsparam1[1])/obsparam1[1]        # RB alpha p 
    rmsep1_iii[s]     <- (p1_iii[s] - obsparam1[1])^2 # RMSE alpha p
    
    rbseff1_iii[s]     <- (bseff1_iii[s]-obsparam1[2])/obsparam1[2]  # RB obs
    rmseseff1_iii[s]   <- (bseff1_iii[s] - obsparam1[2])^2 # RMSE obs
    
    rbp2_iii[s]        <- (p2_iii[s]-obsparam2[1])/obsparam2[1]       # RB alpha p 
    rmsep2_iii[s]      <- (p2_iii[s] - obsparam2[1])^2 # RMSE alpha p
    
    rbseff2_iii[s]     <- (bseff2_iii[s]-obsparam2[2])/obsparam2[2]  # RB obs
    rmseseff2_iii[s]   <- (bseff2_iii[s] - obsparam2[2])^2 # RMSE obs
    
    
    # iv
    rbpsi_iv[s]        <- (psi_iv[s]-covparam[1])/covparam[1]  # RB alpha.psi
    rmsepsi_iv[s]      <- (psi_iv[s] - covparam[1])^2 # RMSE alpha.psi
    
    rbcov_iv[s]        <- (bcov_iv[s]-covparam[2])/covparam[2]   # RB cov
    rmsecov_iv[s]      <- (bcov_iv[s] - covparam[2])^2 # RMSE cov
    
    rbp1_iv[s]         <- (p1_iv[s]-obsparam1[1])/obsparam1[1]        # RB alpha p 
    rmsep1_iv[s]       <- (p1_iv[s] - obsparam1[1])^2 # RMSE alpha p
    
    rbseff1_iv[s]      <- (bseff1_iv[s]-obsparam1[2])/obsparam1[2]   # RB obs
    rmseseff1_iv[s]    <- (bseff1_iv[s] - obsparam1[2])^2 # RMSE obs
    
    rbp2_iv[s]         <- (p2_iv[s]-obsparam2[1])/obsparam2[1]       # RB alpha p 
    rmsep2_iv[s]       <- (p2_iv[s] - obsparam2[2])^2 # RMSE alpha p
    
    rbseff2_iv[s]      <- (bseff2_iv[s]-obsparam2[2])/obsparam2[2]   # RB obs
    rmseseff2_iv[s]    <- (bseff2_iv[s] - obsparam2[2])^2 # RMSE obs
    
    # ibis  
    rbpsi_ibis[s]        <- (psi_ibis[s]-covparam[1])/covparam[1]   # RB alpha.psi
    rmsepsi_ibis[s]      <- (psi_ibis[s] - covparam[1])^2 # RMSE alpha.psi
    
    rbcov_ibis[s]        <- (bcov_ibis[s]-covparam[2])/covparam[2]   # RB cov
    rmsecov_ibis[s]      <- (bcov_ibis[s] - covparam[2])^2 # RMSE cov
    
    rbp_ibis[s]          <- (p_ibis[s]-obsparam2[1])/obsparam2[1]       # RB alpha p 
    rmsep_ibis[s]        <- (p_ibis[s] - obsparam2[1])^2 # RMSE alpha p
    
    rbseff_ibis[s]       <- (bseff_ibis[s]-obsparam2[2])/obsparam2[2]   # RB obs
    rmseseff_ibis[s]     <- (bseff_ibis[s] - obsparam2[2])^2 # RMSE obs
    
    
    # ii bis
    rbpsi_iibis[s]       <- (psi_iibis[s]-covparam[1])/covparam[1]   # RB alpha.psi
    rmsepsi_iibis[s]     <- (psi_iibis[s] - covparam[1])^2 # RMSE alpha.psi
    
    rbcov_iibis[s]       <- (bcov_iibis[s]-covparam[2])/covparam[2]  # RB cov
    rmsecov_iibis[s]     <- (bcov_iibis[s] - covparam[2])^2 # RMSE cov
    
    rbp_iibis[s]         <- (p_iibis[s]-obsparam2[1])/obsparam2[1]      # RB alpha p 
    rmsep_iibis[s]       <- (p_iibis[s] - obsparam2[1])^2 # RMSE alpha p
    
    rbseff_iibis[s]      <- (bseff_iibis[s]-obsparam2[2])/obsparam2[2]   # RB obs
    rmseseff_iibis[s]    <- (bseff_iibis[s] - obsparam2[2])^2 # RMSE obs
    
    
  } # end for 
  
  

  para <- data.frame("nsites" = nsites,"nocc" =  nocc, "nsim" = nsim, "covparam"= covparam, "obsparam1"= obsparam1,"obsparam2" = obsparam2)
  
  return(list(para, # i 
              rbpsi_i,rmsepsi_i , rbcov_i ,
              rmsecov_i,
              rbp_i ,
              rmsep_i ,
              rbseff_i , 
              rmseseff_i ,
              # ii
              rbpsi_ii,
              rmsepsi_ii,
              rbcov_ii,
              rmsecov_ii,
              rbp_ii,
              rmsep_ii,
              rbseff_ii,
              rmseseff_ii ,
              # iii
              rbpsi_iii,
              rmsepsi_iii ,
              rbcov_iii,
              rmsecov_iii ,
              rbp1_iii,
              rmsep1_iii,
              rbseff1_iii ,
              rmseseff1_iii,
              rbp2_iii,
              rmsep2_iii,
              rbseff2_iii ,
              rmseseff2_iii,
              # iv
              rbpsi_iv ,
              rmsepsi_iv,
              rbcov_iv,
              rmsecov_iv,
              rbp1_iv,
              rmsep1_iv,
              rbseff1_iv,
              rmseseff1_iv,
              rbp2_iv ,
              rmsep2_iv,
              rbseff2_iv,
              rmseseff2_iv, 
              # ibis
              rbpsi_ibis,rmsepsi_ibis , rbcov_ibis ,
              rmsecov_ibis,
              rbp_ibis ,
              rmsep_ibis ,
              rbseff_ibis , 
              rmseseff_ibis ,
              # iibis
              rbpsi_iibis,
              rmsepsi_iibis,
              rbcov_iibis,
              rmsecov_iibis,
              rbp_iibis,
              rmsep_iibis,
              rbseff_iibis,
              rmseseff_iibis))
  
} # endfunction

```

R code to run the analysis  

```{r eval=FALSE}
nsim = 100

nsites = 100
nocc = 4

covparam.3 <- c(-0.5,0.2,0) # psi = 0.3
covparam.1 <- c(-1.9,0.2,0) # psi = 0.1

covparam <- list(covparam.1 = c(-1.9,0.2,0),covparam.3 = c(-0.5,0.2,0))

obsparam <- c(-0.1,0.16,0) # p = 0.5
obsparam1 <- c(-0.1,0.16,0) # p = 0.5

obsparam2 <- c(1.8,0.6,0) # p = 0.8

# all simul for loop 
res <- list()
i <- 0
t <- proc.time()
for(c in 1: length(covparam)){  #i <- i+1
  

    i <- i+1 
    
    print(paste("obs",o))
    print(paste("cov",c))
    res[[i]] <- loop(nsites, nocc, nsim, covparam[[c]], obsparam1,obsparam2, Seff, Seff2, Covenv)
    
    
} # end c

run.time <- proc.time() - t


names(res[[1]]) <-  names(res[[2]]) <-  c("para", "rbpsi_i","rmsepsi_i" , "rbcov_i" ,
  "rmsecov_i",
  "rbp_i" ,
  "rmsep_i" ,
  "rbseff_i" , 
  "rmseseff_i" ,
  "rbpsi_ii",
  "rmsepsi_ii",
  "rbcov_ii",
  "rmsecov_ii",
  "rbp_ii",
  "rmsep_ii",
  "rbseff_ii",
  "rmseseff_ii" ,
  "rbpsi_iii",
  "rmsepsi_iii" ,
  "rbcov_iii",
  "rmsecov_iii" ,
  "rbp1_iii",
  "rmsep1_iii",
  "rbseff1_iii" ,
  "rmseseff1_iii",
  "rbp2_iii",
  "rmsep2_iii",
  "rbseff2_iii" ,
  "rmseseff2_iii",
  "rbpsi_iv",
  "rmsepsi_iv",
  "rbcov_iv",
  "rmsecov_iv",
  "rbp1_iv",
  "rmsep1_iv",
  "rbseff1_iv",
  "rmseseff1_iv",
  "rbp2_iv" ,
  "rmsep2_iv",
  "rbseff2_iv",
  "rmseseff2_iv", 
  # bis
  "rbpsi_ibis","rmsepsi_ibis" , "rbcov_ibis" ,
  "rmsecov_ibis",
  "rbp_ibis" ,
  "rmsep_ibis" ,
  "rbseff_ibis" , 
  "rmseseff_ibis" ,
  "rbpsi_iibis",
  "rmsepsi_iibis",
  "rbcov_iibis",
  "rmsecov_iibis",
  "rbp_iibis",
  "rmsep_iibis",
  "rbseff_iibis",
  "rmseseff_iibis")
 
  names(res) <- c("occ1", "occ3")
  
save(res,run.time,file = "simIOM.rdata")
```


# Results

<center>
```{r, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE,fig.fullwidth = TRUE, fig.height= 8, fig.width=16 , fig.cap="Fig. 3: RMSE and RB for covariate effect size, and predicted occupancy probability"}
 
  nsim = 100
  
  load("simIOM.rdata")
  
  # make a tibble to store values 
  extract.RMSE <- function(res, para){
    m <- c()
    for(i in 1:length(res)){
      m[i] <- sqrt(mean(res[[i]][[para]]))
    } # i
    return(m)
  }
  
  extract.RB <- function(res, para){
    m <- c()
    for(i in 1:length(res)){
      m[i] <- mean(res[[i]][[para]])
    } # i
    return(m)
  }
  

  res1 <- res$occ1
  res2 <- res$occ3
  
  rocc <- tibble(model = rep(c("cov_i","cov_ii","cov_ibis","cov_iibis","cov_iii","cov_iv"), 2),
               rb.occ = c(mean(res1$rbcov_i),
                          mean(res1$rbcov_ii),
                          mean(res1$rbcov_ibis),
                          mean(res1$rbcov_iibis),
                          mean(res1$rbcov_iii),
                          mean(res1$rbcov_iv),
                          mean(res2$rbcov_i),
                          mean(res2$rbcov_ii),
                          mean(res2$rbcov_ibis),
                          mean(res2$rbcov_iibis),
                          mean(res2$rbcov_iii),
                          mean(res2$rbcov_iv)), 
               rmse.occ = c(sqrt(mean(res1$rmsecov_i)),
                          sqrt(mean(res1$rmsecov_ii)),
                          sqrt(mean(res1$rmsecov_ibis)),
                          sqrt(mean(res1$rmsecov_iibis)),
                          sqrt(mean(res1$rmsecov_iii)),
                          sqrt(mean(res1$rmsecov_iv)),
                          sqrt(mean(res2$rmsecov_i)),
                          sqrt(mean(res2$rmsecov_ii)),
                          sqrt(mean(res2$rmsecov_ibis)),
                          sqrt(mean(res2$rmsecov_iibis)),
                          sqrt(mean(res2$rmsecov_iii)),
                          sqrt(mean(res2$rmsecov_iv))))
    
    rocc <- rocc %>% mutate(type = rep(c("RV","SV"),length(model)/2),
           occ = c(rep("Ψ = 0.1",length(model)/2),rep("Ψ = 0.3", length(model)/2)),
           name = rep(c("SOM p=0.1","SOM p=0.1","SOM p=0.5","SOM p=0.5","IOM","IOM"),2))
  
  
  pocc1 <- rocc %>%
    ggplot() +
    aes(x = name, y = rmse.occ, fill = type) + 
    geom_col(position = "dodge", width=.6) + 
    labs(x = '',
         y = 'RMSE in occupancy covariate effect size') + 
    scale_fill_manual(name = NULL, values=c("#000A39","#d2a34e"))+ 
    facet_wrap(~ occ,  strip.position = "top", labeller = label_wrap_gen(multi_line = FALSE)) + 
    theme_bw(base_size = 14)
  #pocc1
  
  pocc2 <- rocc %>%
    ggplot() +
    aes(x = name, y = rb.occ, fill = type) + 
    geom_col(position = "dodge", width=.6) + 
    labs(x = '',
         y = 'RB in occupancy covariate effect size') + 
    ylim(-2.5,1.5)+
    scale_fill_manual(name = NULL, values=c("#000A39","#d2a34e"))+ 
    facet_wrap(~ occ,  strip.position = "top", labeller = label_wrap_gen(multi_line = FALSE)) + 
    theme_bw(base_size = 14)
  

  #--------- psi reconstructed -----
  
#names(res1)

  # extract parameter
  nsites <- 100
 covenv <- runif(nsites,0,1)
  
  psi.rv1 <-c()
  psi.sv1 <-c()
  psi.rv2 <-c()
  psi.sv2 <-c()
  psi.rviom <-c()
  psi.sviom <-c()
  psi.t <- c()
  
  rb.psi.rv1  <- c()
  rmse.psi.rv1 <- c()
  rb.psi.sv1  <- c()
  rmse.psi.sv1 <- c()
  
  rb.psi.rv2  <- c()
  rmse.psi.rv2 <- c()
  rb.psi.sv2  <- c()
  rmse.psi.sv2 <- c()
  
  rb.psi.rviom  <- c()
  rmse.psi.rviom <- c()
  rb.psi.sviom  <- c()
  rmse.psi.sviom <- c()
  
 m <- list()
# ---- res1 ----
    m[[1]] <- res1$rbpsi_i*  res1$para$covparam[1] +  res1$para$covparam[1]  #Psi RV1
    m[[2]] <- res1$rbcov_i*  res1$para$covparam[2]  + res1$para$covparam[2] # cov Psi RV1
    m[[3]] <- res1$rbpsi_ii*  res1$para$covparam[1] + res1$para$covparam[1] # int psi SV1
    m[[4]] <- res1$rmsecov_ii*res1$para$covparam[2]  +res1$para$covparam[2]  # cov SV1
    m[[5]] <- res1$rbpsi_ibis*res1$para$covparam[1]  +res1$para$covparam[1]  # int psi RV2
    m[[6]] <- res1$rbcov_ibis*res1$para$covparam[2]  +res1$para$covparam[2]  # cov psi RV2
    m[[7]] <- res1$rbpsi_iibis* res1$para$covparam[1] + res1$para$covparam[1] # int psi SV2
    m[[8]] <- res1$rbcov_iibis* res1$para$covparam[2] + res1$para$covparam[2] # cov SV2
    m[[9]] <- res1$rbpsi_iii*res1$para$covparam[1]  +res1$para$covparam[1]  # int psi IOM RV
    m[[10]] <- res1$rbcov_iii*res1$para$covparam[2]  +res1$para$covparam[2]  # cov psi IOM RV
    m[[11]] <- res1$rbpsi_iv* res1$para$covparam[1] + res1$para$covparam[1] # int psi IOM SV
    m[[12]] <- res1$rbcov_iv* res1$para$covparam[2] + res1$para$covparam[2] # cov IOM SV
    
    psi.rv1 <- 1/(1+exp(- m[[1]] - m[[2]]*scale(covenv)))
    psi.sv1 <- 1/(1+exp(- m[[3]] - m[[4]]*scale(covenv)))
    psi.rv2 <- 1/(1+exp(- m[[5]] - m[[6]]*scale(covenv)))
    psi.sv2 <- 1/(1+exp(- m[[7]] - m[[8]]*scale(covenv)))
    psi.rviom <- 1/(1+exp(- m[[9]] - m[[10]]*scale(covenv)))
    psi.sviom <- 1/(1+exp(- m[[11]] - m[[12]]*scale(covenv)))
    
psi.t1 <- 1/(1+exp(- res1$para$covparam[1] - res1$para$covparam[2]*scale(covenv)))

    Mrb.psi.rv1  <-  mean((psi.rv1 - psi.t1)/psi.t1)
    Mrmse.psi.rv1 <- sqrt(mean((psi.rv1 - psi.t1)^2))
    Mrb.psi.sv1  <- mean((psi.sv1 - psi.t1)/psi.t1)
    Mrmse.psi.sv1 <- sqrt(mean((psi.sv1 - psi.t1)^2))
    Mrb.psi.rv2  <- mean((psi.rv2 - psi.t1)/psi.t1)
    Mrmse.psi.rv2 <- sqrt(mean((psi.rv2 - psi.t1)^2))
    Mrb.psi.sv2  <- mean((psi.sv2 - psi.t1)/psi.t1)
    Mrmse.psi.sv2 <- sqrt(mean((psi.sv2 - psi.t1)^2))
    Mrb.psi.rviom  <- mean((psi.rviom - psi.t1)/psi.t1)
    Mrmse.psi.rviom <- sqrt(mean((psi.rviom - psi.t1)^2))
    Mrb.psi.sviom  <- mean((psi.sviom - psi.t1)/psi.t1)
    Mrmse.psi.sviom <- sqrt(mean((psi.sviom - psi.t1)^2))

    m <- list()
    # ---- res2-----  
    m[[1]] <- res2$rbpsi_i*     res2$para$covparam[1] +   res2$para$covparam[1]  #Psi RV1
    m[[2]] <- res2$rbcov_i*     res2$para$covparam[2]  +  res2$para$covparam[2] # cov Psi RV1
    m[[3]] <- res2$rbpsi_ii*    res2$para$covparam[1] +   res2$para$covparam[1] # int psi SV1
    m[[4]] <- res2$rmsecov_ii*  res2$para$covparam[2]  +  res2$para$covparam[2]  # cov SV1
    m[[5]] <- res2$rbpsi_ibis*  res2$para$covparam[1]  +  res2$para$covparam[1]  # int psi RV2
    m[[6]] <- res2$rbcov_ibis*  res2$para$covparam[2]  +  res2$para$covparam[2]  # cov psi RV2
    m[[7]] <- res2$rbpsi_iibis* res2$para$covparam[1] +   res2$para$covparam[1] # int psi SV2
    m[[8]] <- res2$rbcov_iibis* res2$para$covparam[2] +   res2$para$covparam[2] # cov SV2
    m[[9]] <- res2$rbpsi_iii*   res2$para$covparam[1]  +  res2$para$covparam[1]  # int psi IOM RV
    m[[10]] <- res2$rbcov_iii*  res2$para$covparam[2]  +  res2$para$covparam[2]  # cov psi IOM RV
    m[[11]] <- res2$rbpsi_iv*   res2$para$covparam[1] +   res2$para$covparam[1] # int psi IOM SV
    m[[12]] <- res2$rbcov_iv*   res2$para$covparam[2] +   res2$para$covparam[2] # cov IOM SV
    
    psi.rv1 <- 1/(1+exp(- m[[1]] - m[[2]]*scale(covenv)))
    psi.sv1 <- 1/(1+exp(- m[[3]] - m[[4]]*scale(covenv)))
    psi.rv2 <- 1/(1+exp(- m[[5]] - m[[6]]*scale(covenv)))
    psi.sv2 <- 1/(1+exp(- m[[7]] - m[[8]]*scale(covenv)))
    psi.rviom <- 1/(1+exp(- m[[9]] - m[[10]]*scale(covenv)))
    psi.sviom <- 1/(1+exp(- m[[11]] - m[[12]]*scale(covenv)))
    
    psi.t2 <- 1/(1+exp(- res2$para$covparam[1] - res2$para$covparam[2]*scale(covenv)))
    
    Nrb.psi.rv1  <-  mean((psi.rv1 - psi.t2)/psi.t2)
    Nrmse.psi.rv1 <- sqrt(mean((psi.rv1 - psi.t2)^2))
    Nrb.psi.sv1  <- mean((psi.sv1 - psi.t2)/psi.t2)
    Nrmse.psi.sv1 <- sqrt(mean((psi.sv1 - psi.t2)^2))
    Nrb.psi.rv2  <- mean((psi.rv2 - psi.t2)/psi.t2)
    Nrmse.psi.rv2 <- sqrt(mean((psi.rv2 - psi.t2)^2))
    Nrb.psi.sv2  <- mean((psi.sv2 - psi.t2)/psi.t2)
    Nrmse.psi.sv2 <- sqrt(mean((psi.sv2 - psi.t2)^2))
    Nrb.psi.rviom  <- mean((psi.rviom - psi.t2)/psi.t2)
    Nrmse.psi.rviom <- sqrt(mean((psi.rviom - psi.t2)^2))
    Nrb.psi.sviom  <- mean((psi.sviom - psi.t2)/psi.t2)
    Nrmse.psi.sviom <- sqrt(mean((psi.sviom - psi.t2)^2))
  
  # plot 
  r <- tibble(model = rocc$name[1:(nrow(rocc))],
              occ = rocc$occ,
              type= rocc$type,
              rmse.psi = c(Mrmse.psi.rv1,Mrmse.psi.sv1,Mrmse.psi.rv2,Mrmse.psi.sv2,Mrmse.psi.rviom,Mrmse.psi.sviom,
                           Nrmse.psi.rv1,Nrmse.psi.sv1,Nrmse.psi.rv2,Nrmse.psi.sv2,Nrmse.psi.rviom,Nrmse.psi.sviom), 
              rb.psi = c(Mrb.psi.rv1,Mrb.psi.sv1,Mrb.psi.rv2,Mrb.psi.sv2,Mrb.psi.rviom,Mrb.psi.sviom,
                         Nrb.psi.rv1,Nrb.psi.sv1,Nrb.psi.rv2,Nrb.psi.sv2,Nrb.psi.rviom,Nrb.psi.sviom))
  
  
  
  ppsi1 <- r %>%
    ggplot() +
    aes(x = model, y = rmse.psi, fill = type) + 
    geom_col(position = "dodge", width=.6) + 
    labs(x = '',
         y = 'RMSE in Ψ') + 
    scale_fill_manual(name = NULL, values=c("#000A39","#d2a34e"))+ 
    facet_wrap(~ occ,  strip.position = "top", labeller = label_wrap_gen(multi_line = FALSE)) + 
    theme_bw(base_size = 14)
  
  ppsi2 <- r %>%
    ggplot() +
    aes(x = model, y = rb.psi, fill = type) + 
    geom_col(position = "dodge", width=.6) + 
    labs(x = '',
         y = 'RB in occupancy Ψ') + 
    ylim(-2.5,1.5)+
    scale_fill_manual(name = NULL, values=c("#000A39","#d2a34e"))+ 
    facet_wrap(~ occ,  strip.position = "top", labeller = label_wrap_gen(multi_line = FALSE)) + 
    theme_bw(base_size = 14)
  
  plot_grid(pocc1,pocc2, ppsi1, ppsi2,ncol = 2, nrow = 2, labels = "AUTO", label_size = 15)
  
```
</center>


## Covariate effect size on occupancy probability 

Concerning covariate effect size on $\psi$ (Fig. 3A-B), the results were similar whatever the occupancy model we considered. We found a better precision and bias for higher occupancy ($\psi = 0.5$ and $\psi = 0.9$). Results between single- and repeated-visits are closed to each other. We also found a slightly lower RMSE and RB for Integrated occupancy models in every scenario.


## Predicted occupancy probabiltiy 

When reconstructing the occupancy probability $\psi$ in Fig. 3C-D, integrated occupancy models exhibit better precision and bias than single-method occupancy models. Both single- and repeated-visits occupancy models seem to benefit from data integration in their performances. Note that RMSE plotting scale is different between Fig. 3A and 3C. 


# Discussion

Our simulation study showed that integrated occupancy models improved precision and bias compared to occupancy models using the datasets in isolation. 
To be consistent with our manuscript, we reported performances of covariate effect size and predicted occupancy. To have a complete understanding of models performances, one might want to look at wider range of relationships, and observe all models parameters (e.g. intercept, sampling effort effect size on detection probability). Integrated occupancy, whether it is repeated-visit or single-visit, should help estimating detection probability with better precision.  
We noticed strong RMSE and RB for single-visit model compared to repeated-visit model when considering $\psi = 0.1$ and $p = 0.5$. Both results from @peachSinglevisitDynamicOccupancy2017, and from our simulation study comparing single- and repeated-visit pointed out the limited performance of single-visit in the case of low occupancy. Integrated single-visit occupancy models help improving RMSE and RB.  

Overall, our simulations underline that integrated single-visit occupancy models can be used to obtain reliable estimates of occupancy.

# References